---
title: "Covid91 vaccine study Final2021"
author: "Dom Ciriello"
date: "20210505"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods  

The first step is isolating the 4 subgroups/sub-populations. After isolating 4 subgroups I will be evaluating the efficacy of the drug for each subgroup.
```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex <- as.factor(maledata$sex)
```

# Males

## Bar Chart  

```{r}
Males <- maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that 58% of the infected males were given the drug and 42% were given the placebo.

## Numerical Summary  

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

Based on the numerical summary, 337 of the 10,400 males in the study were infected with COVID. Of those 337, 197 had received the vaccine and only 141 received the placebo. The table shows that 3.72% of the males who received the vaccine contracted COVID. 2.55% of the males who received the placebo contracted COVID. The vaccine has little to no effect. However, the small effect that it did have is negative. A higher number/percentage of the infected males received the vaccine. Males who got the vaccine were about 50% more likely to contract COVID.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Since the p-value is extremely small, the odds of our results happening purely by chance is extremely unlikely.  

$H_0$  
The null hypothesis is that this vaccine makes no difference in whether or not you contract COVID for males.  

$H_A$  
The alternative hypothesis is that the vaccine does make a difference for males in whether or not they contract COVID.  

Therefore we reject the null hypothesis.  

You can convert our odds ratio of 1.47 into an effectiveness percentage by dividing the odds ratio by 1+odds ratio and then multiplying by 100.  

100(1.47/(1+1.47)) = 59.5% effective

# Females

# LGBTQRSTUVWXYZ (Alphabet Individuals)

# Drug Users

# Discussion & Conclusion